The Analog Method for Precipitation Prediction: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-24 Moving temporal window - HESS/template.tex
New version of the introduction.

\title{The Analog Method for Precipitation Downscaling: Finding Better Analog Situations at a Sub-Daily Time Step}

@@

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

@@

However, attempts to reconstruct longer precipitation series of running 24~h-totals by means of simple methods failed. It emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of the statistical downscaling methods \cite{Maraun2010}. They consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Bontron2005}. Multiple variations of the methods exist \citep[a non exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

AMs are usually implemented with a daily time step (see references above), due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analog situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day. Previous tests showed the benefit, in terms of analogy criterion values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the prediction skills \citep{Finet2008}.

The time step at which the MTW is applied depends on the time step of the predictor variables. In the present study, a 6-hourly reanalysis dataset is used, which is a standard temporal resolution. Thus, additionally to eventually finding better analogs, the MTW also increases the archive length. However, despite having four times more candidate situations, the quantity of new information is not expected to be as important, due to significant autocorrelation between successive situations at a 6-hourly time step. The quality of the analogy depends on the archive length to a certain extent \cite{Ruosteenoja1988, Vandendool1994}. Thus, the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, whereas the archive inflation brought by the MTW is expected to contribute to an increase in performance.

@@

@@

Applying the MTW technique involve the need of sub-daily precipitation timeseries. However, such data are not available on long periods. As the AM benefits from long archives, it would be profitable to be able to apply the MTW to a longer archive. Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily time series were assessed.

@@

The second approach aims at getting closer to the chronology of the actual precipitation by relying on some informative variable during the reconstruction procedure. The best proxy variable is identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution.

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. This had to be assessed and the MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). The improvement of the performance is not very large, but is nevertheless substantial. Moreover, it requires no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated.

@@

@@

A higher number of analogs generally means, with an archive of fixed length, a poorer analogy. Indeed, when the choice of the predictors or the parameters are improved, leading to a better prediction, the optimal number of analog situations decreases. However, when the length of the archive increases, the optimal number of analogs increases too, for a better performance, up to a certain threshold (\citealt{Bontron2004}, see also Fig. 7 of \citealt{Hamill2006a}). The observed increases in the number of analog situations with the MTW result in better performance skills, as it can be seen as an inflation of the archive. The MTW enriches the pool of available situations, even though they are not fully independent.

@@

@@

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. First, these archives of high temporal resolution precipitation data are increasing now with time. Another possible source of such archives is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. Even though outputs from these models might be biased or not accurate enough, information on the timing of the precipitation events could be useful in order to disaggregate stations time series. When precipitation data with accurate time distribution will be available, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

NCEP reanalysis data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at http://www.esrl.noaa.gov/psd/. Precipitation time series provided by MeteoSwiss. 
\end{acknowledgements}

>>>>

\title{The Analog Method for Precipitation Prediction: Finding Better Analog Situations at a Sub-Daily Time Step}

@@

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

@@

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM should be considered for any application, and in any context, let it be operational forecasting or climate-related studies

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of the statistical downscaling methods \cite{Maraun2010}. They consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the methods exist \citep[a non exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method was recently analyzed in depth \citep{Chardon2014}. A great advantage of AMs is that they create realistic events, because they were observed, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} have put a major effort into analyzing the influence of the archive length on the quality of the analogy. They have highlighted a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (of degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrated that it is not possible to find good analogues on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to look for analogues, and the archive length is maximized.

Thus, due to the availability of long precipitation archives at a daily time step that have no equivalent at a finer resolution, AMs are usually implemented on a daily basis (see references above). Therefore, the analog situations are assessed by comparing predictors at fixed hours of the day. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day. Previous tests showed the benefit, in terms of analogy criterion values, of searching for analog synoptic situations at a finer time step, but without assessing the impact on the prediction skills \citep{Finet2008}.

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant autocorrelation between successive situations. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does bring some new information from the archive. Thus, the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, whereas the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new options \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

@@

More complex AMs exist, with additional predictors \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. The MTW can also be applied on these, as they generally rely on a similarity of the atmospheric circulation in the first level of analogy, but it is easier to interpret the impact of the MTW on more basic methods.

@@

Applying the MTW technique involve the need of sub-daily precipitation timeseries. However, such data are not available on long periods. As the AM benefits from long archives, it would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008). Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily time series were assessed.

@@

The second approach aims at getting closer to the chronology of the actual precipitation by relying on some informative variable during the reconstruction procedure. The best proxy variable is identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution. Then, once the best proxy has been selected, its temporal profile is used in order to disaggregate the precipitation time series on the long archive period.

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. This had to be assessed and the MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). The improvement of the performance is not very large, but is nevertheless substantial. Moreover, it requires no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated. It is here of the same order of magnitude as if the length of the archive doubled. Note that despite the number of candidate situations are four times higher, the gains seem to be lower than for a quadrupled archive length. The likely reason is that an actual longer archive contains more atmospheric situations that might have been less observed on a shorter period. Moreover, high correlations between sub-daily circulation patterns are expected. The MTW thus does not create four times more independent data, but the increase in information is nevertheless substantial. In this regards, a reanalysis dataset with a higher temporal resolution might not improve the performance significantly than a 6-hourly. On the other hand, reducing the MTW to a 12-hourly time step might reduce redundancy in the archive, but might also reduce the chance of finding better analog situations, as circulation patterns can evolve substantially in 12~hours.

Moreover, in a transient climate, an eventual non-stationarity of the link between predictors and precipitation might discard the relevance of analogs from far in the past, and thus increase the relevance of using a more recent and shorter archive rather than a long one. In such case, the archive inflation brought by the MTW is also relevant.

@@

We just mentioned that the MTW improves the prediction of heavier precipitation events. However, it will have no effect on the prediction of extreme events that were not yet observed, and therefore are not present in the archive. Extremes in AMs can be modeled e.g. by extrapolation of a truncated exponential or gamma distribution fitted on the analog values \citep{Obled2002}. Another possible approach is by combining AMs with other methods \citep[e.g.][]{Chardon2014}. However, this goes beyond the scope of the present study, as prediction of unobserved extremes are not modified by the MTW.

@@

A higher number of analogs generally means, with an archive of fixed length, a poorer analogy. Indeed, when the choice of the predictors or the parameters are improved, leading to a better prediction, the optimal number of analog situations decreases. However, when the length of the archive increases, the optimal number of analogs increases too, for a better performance, up to a certain threshold (\citealt{Bontron2004}, see also Fig. 7 of \citealt{Hamill2006a}). The observed increases in the number of analog situations with the MTW result in better performance skills for a given method, as it can be seen as an inflation of the archive. However, if new relevant predictors were added to the method, the number of analogs would then decrease.

@@

%TODO: : it was shown by some studies that the resolution of the reanalysis dataset does not improve significantly the performance of the analogy of the atmospheric circulation. However, a full analysis of this aspect is out of the scope of this publication, and is a topic we are working on right now. Nevertheless, we can make the assumption that it does not alter the fact that we can find better analog situations at different hours of the day.

@@

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. First, these archives of high temporal resolution precipitation data are increasing now with time. Another possible source of such archives is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. Even though outputs from these models might be biased or not accurate enough, information on the timing of the precipitation events could be useful in order to disaggregate stations time series. When precipitation data with accurate time distribution will be available, the use of a MTW in the AM should be considered for any application, and in any context, let it be operational forecasting or climate-related studies.

@@

NCEP reanalysis data provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado, USA, from their Web site at http://www.esrl.noaa.gov/psd/. Precipitation time series provided by MeteoSwiss.

The authors would also like to acknowledge the work anonymous reviewers, which contributed to improving this paper. 
\end{acknowledgements}
